---
title: "Example Usage of This Package"
author: "Liang Wu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Usage of This Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

This package implement functions `pearson.r`, `rankdata` and `spearman.r` to calculate
correlation coefficient in defferent methods. Here are usage of these functions:

### `pearson.r`

```{r}
library(BA22022002)

xs <- runif(1000)
ys <- 0.2*xs + runif(1000)
pearson.r(xs, ys)
```

Pearson correlation can show the positive correlation between `xs` and `ys`.

### `rankdata`

Spearman correlation is just the same as pearson correlation but calculated 
with rank of original data. So I implement `rankdata` function first, with
`Rcpp`. And here is usage.


```{r}
rankdata(c(1, 5, 2))
```

`rankdata` return rank of input data.


### `spearman.r`

Spearman correlation can be calculated using `pearson.r` and `rankdata`. And here is an example:

```{r}
xs[1] <- -100
ys[1] <- 100
c(pearson.r(xs, ys), spearman.r(xs, ys))
```

Pearson correlation can be strongly affect by outliers, while spearman correlation is more stable.
